Introducing statistics with \\texttt{intRo}

### 2014-12-04 student-paper-competition-2014/paper.Rmd
Draft the abstract, intro, conclusion

\begin{abstract}
Your abstract goes here...
...
\end{abstract}

@@

# Introduction

@@

\ak{[Make sure we talk about the upcoming R package for module creation, mention also having intro in the classroom in the spring]}
\ak{[Future challenges: server performance?]}

>>>>

\begin{abstract}
\texttt{intRo} is a modern web-based application for performing basic data analysis and statistical routines. Leveraging the power of \texttt{R} and \texttt{Shiny}, \texttt{intRo} implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician. This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}. \texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}
\end{abstract}

@@

# Introduction
The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community. However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student. This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work. Still, signficant compromises must be made when using software like \texttt{JM}P, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions. In teaching STAT 226 at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class. Often, these struggles were directly related to the use of the software and not any sort of fundamental misunderstanding of the material.

The release of \texttt{RStudio}'s \texttt{Shiny}, which allows for easily turning an R-based analysis into an interactive web application, led to the creation of an introductory statistics application which we call \texttt{intRo}. \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.

@@

We believe \texttt{intRo} can be a powerful and effective tool for introductory statistics education. Its modular structure allows it to be flexible enough for many different applications and curriculums. Its ease-of-use allows the student to focus his or her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation. Reproducible code generated from each analysis can be used to spark an interest in \texttt{R} programming in those who might otherwise not be exposed to it.

Despite the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more accessible and useful. In particular, we have begun development on an \texttt{R} package which will allow modules to be created automatically. This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo}. This will vastly improve intRo's flexibility and allow it to be used in a wider range of curriculums, including more advanced statistics courses. We will also be trialing \texttt{intRo}'s use in a statistics course at Iowa State University in Spring Semester 2015. This will allow us to assess its usefulness relative to software used in the past, as well as areas for improvement. Challenges exist with regards to the wider adoption of \texttt{intRo}. For instance, we will need to monitor how well the server \texttt{intRo} is hosted on handles the load of dozens of students performing data analyses. If performance issues are noted, the infrastructure used may need to be expanded to handle current and future load.
